---
title: "Assignment 01"
author: "Adithya KR"
date: "2024-02-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Write an R code snippet to calculate the fourth root of the number 9876 and then find its exponential.(1 mark)  

```{r echo=TRUE}
forth_root <- 9876^(1/4)
forth_root
```
```{r echo=TRUE}
exponential <-exp(9876)
exponential
```
### 2.Create a vector in R that follows this pattern: (5, 10, 15, 20, 25, 20, 15, 10, 5), where the numbers start at 5 and increase by 5 up to 25, and then decrease by 5 back down to 5.(1 mark)  

```{r echo=TRUE}
vec1 <- c(seq(from = 5,to = 25, by =5),seq(from = 20,to = 5, by =-5))
vec1
```
### 3. Create a vector in R that contains the elements 10, 20, and 30, repeated in a pattern of 5 occurrences of 10, 7 occurrences of 20, and 9 occurrences of 30.(1 mark)  
```{r echo=TRUE}
vec2 <- c(rep(10,5),rep(20,7),rep(30,9))
vec2
```
\pagebreak
### 4. Write a program in R for a shipping company that calculates the total shipping cost based on the weight of the items. The shipping costs are tiered as follows: (2 marks)  
*  For items weighing up to 10 kg, the shipping cost is 1500.
*  For items weighing more than 10 kg but less than or equal to 20 kg, the shipping cost is 2500.
* For items weighing more than 20 kg, the shipping cost is 4000 plus 500 for each kg over 20.
```{r echo=TRUE}
weight=21
if (weight <=10){
  cost=1500
}else if (weight<=20){
  cost=2500
}else{
  extra=(weight-20)
  cost=4000+(extra*500)
}
cost
```
### 5. You have been given a dataset of the weekly revenue figures for a company over 4 weeks. Write an R program to compute and display the following:(2 marks)
Sample Revenue Data:  
revenue <- c(5000, 7000, 6500, 4800) 

(a) Total revenue for the 4 weeks.
(b) Average weekly revenue.
(c) The week with the maximum revenue.
(d) The week with the minimum revenue.

```{r echo=TRUE}
rev = c(5000, 7000, 6500, 4800)
##Total revenue
sum(rev)
##Average revenue
mean(rev)
##minimum revenue 
min(rev)
##maximum revenue 
max(rev)
```
### 6. Write a program in R that asks the user to enter their age and then provides health recommendations based on the following age groups:(2 marks)  

* Children (up to 12 years): Emphasize outdoor play and balanced diet.
* Teenagers (13-19 years): Focus on regular exercise and avoiding junk food. 
* Adults (20-60 years): Recommend regular health check-ups and stress management.
* Seniors (above 60 years): Suggest low-impact exercises and a diet rich in calcium and vitamins.

```{r ,echo=TRUE}
age = 43
if(age<=12){
  recom = "Emphasize outdoor play and balanced diet."
}else if(13<=age & age<=19){
  recom = "Focus on regular exercise and avoiding junk food."
}else if(20<=age & age<=60){
  recom = "Recommend regular health check-ups and stress management."
}else {
  recom = "Suggest low-impact exercises and a diet rich in calcium and vitamins."
}
print(recom)
```
### 7. Create an R program that takes a vector of daily currency values against the dollar for 10 days. Your program should:

### a. Calculate the percentage fluctuation in the currency value from the start to the end of the month.
```{r, echo = TRUE}
vec = c(80,81,79,81.5,90,99,100,83.5,77,89)
len = (length(vec))
fluc = ((vec[len] - vec[1])/vec[1])*100
print(fluc)
```
### b. Order the daily values in ascending order.
```{r, echo=TRUE}
sorted = sort(vec)
print(sorted)
```
### c. Identify and display the top 5 days with the highest currency value and the bottom 5 days with the lowest value.(2 marks)  
```{r, echo=TRUE}
bot5val = head(sorted,5)
top5val = tail(sorted,5)
top5days = which(vec>max(bot5val))#all values greater than the max value of bottom5
bot5days = which(vec<min(top5val))#all values lesser  than the min value of top5
```
```{r, echo=TRUE}
#highest 5 currency values
top5val
#top 5 days with highest currency values
top5days
#lowest 5 currency values
bot5val
#bottom 5 days with lowest currency values
bot5days
```
### 8. Write an R program that simulates a bank transaction history:(2 Marks)

### a. Create a numeric vector transactions containing a mix of positive and negative values, representing deposits and withdrawals.  
```{r, echo=TRUE}
transc_list = c(100,-20,90,190,-200,-800,1000)
transc_list
```
### b. Calculate the total balance by summing the elements of the transactions vector.  
```{r, echo=TRUE}
total = sum(transc_list)
total
```
### c. Identify the months where the balance was negative.
```{r ,echo=TRUE}
cur_bal = 0
for (i in 1:length(transc_list)){
  cur_bal = cur_bal + transc_list[i]
  if(cur_bal < 0){
    cat("month with negative balance = ",i)
  }
}
```
### d. Create a new vector positive_transactions containing only the positive values from the transactions vector.
```{r, echo=TRUE}
pos_transc = transc_list[transc_list>0]
pos_transc
```

### 9. Write a function which takes a single argument which is a matrix. The function should return a matrix which is the same as the function argument but every odd number is tripled and even number doubled on the given matrix.(1 mark)

```{r, echo=TRUE}
modifyMatrix <-function(input_matrix){
  output_matrix = input_matrix
  output_matrix[sapply(input_matrix, function(x) x %% 2 == 1)] =
    input_matrix[sapply(input_matrix, function(x) x %% 2 == 1)] * 3
  output_matrix[sapply(input_matrix, function(x) x %% 2 == 0)] = 
    input_matrix[sapply(input_matrix, function(x) x %% 2 == 0)] * 2
  return(output_matrix)
  
}
original_matrix = matrix(c(1,2,3,4,5,6,7,8,9),nrow = 3,ncol = 3,byrow = TRUE)
modified_matrix = modifyMatrix(original_matrix)
```

```{r ,echo=TRUE}
#original_matrix
original_matrix
#output matrix
modified_matrix
```
\pagebreak
### 10. For this exercise, you are provided with a vector vec2 containing the numbers 23, 15, 42, 55, 92.(3 marks)
vec2 <- c(23, 15, 42, 55, 92)

### a. Use a for loop to print each number along with its logarithm (base 10) and square root
```{r, echo=TRUE}
vec2 = c(23, 15, 42, 55, 92)
for(i in vec2){
  cat("number: ",i," logarithm: ",log10(i), " square_root: ", sqrt(i), "\n")
}
```
### b. Achieve the same result using a while loop
```{r ,echo=TRUE}
i = 1
while(i<=length(vec2)){
  cat("number: ",vec2[i]," logarithm: ",log10(vec2[i]), " square_root: ", sqrt(vec2[i]), "\n")
  i = i+1
}
```
### c. Achieve the same result without using an explicit loop

```{r ,echo=TRUE}
result = data.frame(number = vec2,logarithm = log(vec2), square_root = sqrt(vec2))
result
```
### 12. Calculate the following by writing code snippets:(2 marks)
* $$\sum_{i=1}^{50}(i^3+4i^2-8i)$$  
* $$\sum_{i=1}^{90}(\frac{2^i}{i^2} + \frac{3^i}{i^3})$$  

```{r ,echo= TRUE}
#equation 1
total_sum = 0
for (i in 1:50){
  total_sum = total_sum + i^3 +4*i^2 - 8*i
}
total_sum
```

```{r, echo=TRUE}
#equation 2
total_sum = 0
for (i in 1:90){
  total_sum = total_sum + (2^i/i^2)+(3^i/i^3)
}
total_sum
```
### 13. Consider the following R code snippet that utilizes a switch statement:(1 marks)
```{r ,echo = TRUE, results=FALSE}
calculate_operation <- function(operation, a, b) {
result <- switch(
operation,
"add" = a + b,
"subtract" = a - b,
"multiply" = a * b,
"divide" = a / b,
stop("Invalid operation")
)
return(result)
}
output <- calculate_operation("multiply", 5, 3)
print(output)
```

### Modify the function to include an additional operation “power”, which calculates a raised to the power of b. Show the modified function and provide an example of how it would be called.  

```{r ,echo = TRUE}
calculate_operation <- function(operation, a, b) {
result <- switch(
operation,
"add" = a + b,
"subtract" = a - b,
"multiply" = a * b,
"divide" = a / b,
"power" = a^b,
stop("Invalid operation")
)
return(result)
}
output <- calculate_operation("power", 5, 3)
print(output)
```
