---
title: "2024MCS110023-Assignment-02"
author: "ADITHYA KR"
date: "04/06/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)    
```

```{r}
library(ggplot2)

```

```{r}
data("airquality")

```



```{r}
# Basic preprocessing
airquality <- na.omit(airquality) # Removing NA values
```

```{r}

head(airquality)
```




**1. Obtain the plot of daily ozone levels by month as shown below:(1 Marks)**


```{r}
ggplot(airquality, aes(x = Day, y = Ozone)) +
        ggtitle("Daily Ozone Levels by Month") +
        labs(x = "Day of the month", y = "Ozone (ppb)") +
        scale_x_continuous(breaks = seq(0, 30, 10))+ facet_wrap(~Month)+geom_line(color="red")

```


**2.Obtain the plot of ozone distribution by month as shown below:(2 Marks)**


```{r}
ggplot(airquality, aes(x=Month, y=Ozone,fill=factor(Month))) +geom_boxplot()+
  labs(title="Ozone Distribution by Month",x = "Month", y = "Ozone(ppb)")

```


**3. Obtain the plot of Ozone Levels vs. Wind Speed as shown below:(2 Marks)**


```{r}
ggplot(airquality, aes(x = Wind, y = Ozone,color = Month)) + geom_point()+scale_x_log10() + 
  labs(title = "Ozone Levels Vs Wind Speed",x = "Wind speed(mph) ",y = "Ozone(pph)")

```


**4. Using the airquality dataset in R, explore and visualize the monthly distribution of wind speeds, highlighting potential variations and outliers from May to September?(2 marks)**


```{r}
ggplot(airquality, aes(x=Month, y=Wind,fill=factor(Month))) +geom_boxplot()  + 
  labs(title="Monthly Distribution of Wind Speed",x="Month", y="Wind Speed mph")

```
```{r}

out<-boxplot.stats(airquality$Wind)$out
boxplot(airquality$Wind,
ylab = "Wind",
main = "wind speeds")
mtext(paste("Outliers: ", paste(out, collapse = ", ")))

```


**5. Obtain the result shown below:(1 Mark)**


```{r}
# Convert Month column to a factor for easier grouping
airquality$Month <- factor(airquality$Month)
```

```{r}
# Calculate mean, median, and standard deviation of ozone and temperature monthly-wise
summary_stats <- aggregate(cbind(Ozone, Temp) ~ Month, data = airquality, FUN = function(x) 
  c(mean = mean(x, na.rm = TRUE), median = median(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE)))
print(summary_stats)
```

```{r}
Month <- c(5,6,7,8,9)
Mean_Ozone <-c(23.61538,29.44444,59.11538,59.96154,31.44828) 
Median_Ozone <- c(18,23,60,52,23) 
SD_Ozone <- c(22.22445,18.20790,31.63584,38.68121,24.14182) 
Mean_Temp <- c(66.73077,78.22222,83.88462,83.96154,76.89655)
Median_Temp <- c(66,77,84,82,76) 
SD_Temp <- c(6.533346,7.838651,4.439161,6.666218,8.503549) 
df1 <- data.frame(Month ,Mean_Ozone, Median_Ozone, SD_Ozone, Mean_Temp, Median_Temp, SD_Temp ) 
df1
```


**6.On days where the Ozone levels are above 100, investigate the relationship between Ozone and Temperature. Create a scatter plot of Ozone vs. Temperature for days with Ozone levels above 100.(2 marks)**



```{r}
high_ozone_days <- airquality[airquality$Ozone > 100, ]
x<-high_ozone_days$Temp
y<-high_ozone_days$Ozone
ggplot(high_ozone_days,aes(x=Temp,y=Ozone))+geom_point(color="Blue")+labs(x="Temperature",y="O
zone Level",title = "Scatter Plot of Ozone vs. Temperature ", subtitle="For Days with Ozone Levels above 100")
     
```

**7. Determine the month with the highest average temperature using the airquality dataset in R, and display only that month along with its average temperature?(1 Mark)**


```{r}
data(airquality)

# Aggregate the data by month and calculate the average temperature for each month
monthly_avg_temp <- aggregate(Temp ~ Month, data = airquality, FUN = mean)

# Find the row with the highest average temperature
max_temp_row <- which.max(monthly_avg_temp$Temp)

# Extract the month with the highest average temperature and its corresponding temperature
highest_avg_temp_month <- monthly_avg_temp$Month[max_temp_row]
highest_avg_temp_value <- monthly_avg_temp$Temp[max_temp_row]

# Display the month with the highest average temperature and its average temperature
cat("Month with the highest average temperature:", highest_avg_temp_month, "\n")
cat("Average temperature:", round(highest_avg_temp_value, 2))

```


**8. What is the correlation between Solar Radiation and Ozone concentration?(1 Mark)**



```{r}

airquality_clean <- na.omit(airquality[, c("Solar.R", "Ozone")])


correlation <- cor(airquality_clean$Solar.R, airquality_clean$Ozone)


print(paste("Correlation between Solar Radiation and Ozone concentration:", round(correlation, 2)))

```

**9. Obtain the plot of Monthly Distribution of Wind Speeds in New York as shown below:(3 Marks)**



```{r}

ggplot(airquality, aes(reorder(Month,Wind,FUN=median), y=Wind, fill=factor(Month))) +geom_boxplot()+ labs(title="Monthly Distribution of Wind Speeds in NewYork",x = "Month", y = 
 "Wind Speed (mph)")+geom_jitter(size=0.4, alpha=0.7)
```

**10. Create a new data frame containing the average Ozone concentration for each month. Arrange the months in descending order based on the average Ozone concentration.(1 Mark)**


```{r}
# Load the airquality dataset
data(airquality)

# Aggregate the data by month and calculate the average Ozone concentration for each month
monthly_avg_ozone <- airquality %>%
  group_by(Month) %>%
  summarise(avg_ozone_concentration = mean(Ozone, na.rm = TRUE))

# Arrange the months in descending order based on the average Ozone concentration
monthly_avg_ozone <- monthly_avg_ozone[order(-monthly_avg_ozone$avg_ozone_concentration), ]

# Print the resulting data frame
print(monthly_avg_ozone)
```


**11. Obtain the plot of Distribution of Solar Radiation by Month as shown below:(2 Marks)**



```{r}
data(airquality)

# Convert the 'Month' column to a factor with appropriate levels

airquality$Month <- factor(airquality$Month, levels = 1:12,
                           labels = month.abb)
# Create the plot with boxplot and jittered points
ggplot(airquality, aes(x = Month, y = Solar.R)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  geom_jitter(aes(color = Month), width = 0.2, alpha = 0.5) +
  labs(title = "Distribution of Solar Radiation by Month",
       x = "Month", y = "Solar Radiation") +
  theme_minimal()+scale_color_discrete(name = "Month") 


```


**12.Obtain the plot of Distribution of Wind Speeds by Month as shown below:(2 Marks)**



```{r}
ggplot(airquality,aes(x=Wind))+geom_histogram(aes(y = ..density..),binwidth = 0.5,    
fill="Green", color="black")+facet_wrap(~Month)+geom_density(lwd = 1.2,linetype = 2,  
colour = 6)+labs(title="Distribution of Wind speeds by Month",x="Wind Speed",y="Density")                                                                                        
```











