---
title: "R Markdown-Lab Exercise"
author: "Adithya KR"
date: "01/28/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

RMarkdown is a powerful tool for data analysis and reporting in R. It allows you to create dynamic documents
that combine code, rendered output (like plots), and narrative text.

## Benefits of R Markdown
* __Reproducibility__: By including R code within the document, the analysis can be easily reproduced.
* __Dynamic Reporting__: Update your data or analysis, and the results in the document update automatically.
* __Multiple Formats__: RMarkdown documents can be converted to a variety of formats including HTML,
PDF, and Word

## Example: Data Analysis with the mtcars Dataset
Let’s demonstrate some basic functionalities of RMarkdown using the mtcars dataset.


### Loading Data
data(mtcars)

## Summary Statistics
Here’s a quick look at the summary statistics of the <span style="font-size: 12px;">`mtcars`</span> dataset.
```{r mtcars}
summary(mtcars)
```

## Plotting Data
Creating a plot to visualize the data is straightforward. For example, we can create a scatter plot of mpg
(Miles/(US) gallon) against wt (Weight (1000 lbs)).  


```{r}
plot(mtcars$wt, mtcars$mpg, main="Miles per Gallon vs Car Weight",
     xlab="Weight (1000 lbs)", ylab="Miles/(US) gallon")
```
\pagebreak

## Including Tables
This is an example of including tables in RMarkdown using the mtcars dataset.

```{r echo = FALSE}
library(knitr)
kable(mtcars[1:6,],caption = "Table 1: A subset of the mtcars dataset")
```

For example, __bold text__ and _italic text_ are easy to add. You can also create lists:  

* Item 1
* Item 2
* Item 3

Or numbered lists:  

1. First number
2. Second number
3. Third number



