---
title: "Type of Visualizations Exercise"
author: "Adithya KR"
date: "3/18/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Dataset: Gapminder
```{r echo=TRUE}
install.packages("dplyr")
library(dplyr)
install.packages("ggplot2")
library(ggplot2)
install.packages("gapminder")
library(gapminder)
```
\pagebreak 

### 1. Create a line plot showing the median GDP per capita over time.
```{r echo=TRUE}
by_year <- gapminder %>%
    group_by(year) %>%
    summarize(medianGdpPercap = median(gdpPercap))
ggplot(by_year, aes(x = year, y = medianGdpPercap)) +
    geom_line()+geom_point()+labs(title="Median GDP per capita over time")
```
\pagebreak 

### 2. Create a line plot showing the change in median GDP per capita by continent over time.
```{r echo=TRUE}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color = continent)) +
  geom_line()+geom_point()+labs(title="Median GDP per capita by continent over time")

```
\pagebreak 

### 3. Create a bar plot showing the median GDP per capita for each continent in the year 1952.
```{r echo=TRUE}
by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))
ggplot(by_continent, aes(x = continent, y = medianGdpPercap,fill=continent)) +
  geom_col()+labs(title="Median GDP per capita for each continent 1952")

```
\pagebreak 

### 4. Filter for observations in the Oceania continent in the year 1952. Create a bar plot comparing the GDP per capita between Australia and New Zealand.
```{r echo=TRUE}
oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year == 1952)
ggplot(oceania_1952, aes(x = country, y = gdpPercap,fill=country)) +
  geom_col()+labs(title="GDP
per capita between Australia and New Zealand")
```
\pagebreak 

### 5. Create a histogram showing the distribution of country populations (in millions) in the year 1952.
```{r echo=TRUE}
gapminder_1952 <- gapminder %>%
filter(year == 1952)
ggplot(gapminder_1952, aes(x = pop/1e6)) +
geom_histogram(binwidth = 10,color="blue")+labs(title = "Distribution of country populations 
                                                (in millons) in the year 1952")
```
\pagebreak 

### 6. Create a histogram similar to the previous one but with the x-axis (population) on a logarithmic scale.
```{r echo=TRUE}
gapminder_1952 <- gapminder %>%
filter(year == 1952)
ggplot(gapminder_1952, aes(x = pop /1e6)) +
geom_histogram(fill="orange") +
scale_x_log10()
```
\pagebreak 

### 7. Create a boxplot comparing the distribution of GDP per capita among continents in the year 1952
```{r echo=TRUE}
gapminder_1952 <- gapminder %>%
filter(year == 1952)
ggplot(gapminder_1952, aes(x = continent, y = gdpPercap)) +
geom_boxplot() +
scale_y_log10()

```

